---
editor_options:
  chunk_output_type: console
---

# INTERNAL

## Package load/unload

```{r}
.onLoad <- function(libname, pkgname) {
  
  # clear pkgpins cache
  tryCatch(expr = pkgpins::clear_cache(board = pkgpins::board(pkg = pkgname),
                                       max_age = pal::pkg_config_val(key = "global_max_cache_age",
                                                                     pkg = pkgname)),
           error = function(e) cli::cli_alert_warning(text = "Failed to clear pkgpins cache on load of {.pkg {pkgname}}. Error message: {e$message}"))
}
```

## Avoid `R CMD check` notes about undefined global objects used in magrittr pipes

cf. <https://github.com/tidyverse/magrittr/issues/29#issuecomment-74313262>

```{r}
utils::globalVariables(names = c(".",
                                 # tidyselect fns
                                 "everything",
                                 # other
                                 "is_pro",
                                 "key",
                                 "last_modified",
                                 "name"))
```

## Constansts

```{r}
this_pkg <- utils::packageName()
```

# EXPORTED

## `is_uptodate`

```{r}
#' Test if RStudio is up to date
#'
#' @inheritParams latest_version
#' @param stable Set to `FALSE` in order to test against the latest [RStudio preview build](https://rstudio.com/products/rstudio/download/preview/) instead
#'   of the latest [stable build](https://rstudio.com/products/rstudio/download/).
#'
#' @return `TRUE` if the currently running RStudio version is the latest version, `FALSE` otherwise.
#' @export
is_uptodate <- function(stable = TRUE,
                        os = NULL) {
  
  rstudioapi::versionInfo()$version >= latest_version(type = rstudioapi::versionInfo()$mode,
                                                      stable = stable,
                                                      os = os)
}
```

## `latest_version`

**Remarks:**

-   cf. <https://askubuntu.com/questions/1237088/automatizing-rstudio-updates>

-   RStudio seems to host binary releases on three different domains:

    -   [RStudio Desktop](https://rstudio.com/products/rstudio/#rstudio-desktop): <https://download1.rstudio.org/>
    -   [RStudio Server](https://rstudio.com/products/rstudio/#rstudio-server): <https://download2.rstudio.org/>
    -   [RStudio Desktop/Server Preview](https://rstudio.com/products/rstudio/download/preview/) and [RStudio Desktop/Server daily
        builds](https://dailies.rstudio.com/): <https://s3.amazonaws.com/rstudio-ide-build/>

-   On each stable release subdomain there's a file named `current.ver` which stores the very latest version number across all releases. But since it doesn't
    differentiate between the open-source and the proprietary RStudio editions (e.g. [RStudio Desktop](https://rstudio.com/products/rstudio/#rstudio-desktop)
    vs. [RStudio Desktop Pro](https://rstudio.com/products/rstudio-desktop-pro/download-commercial/)) and the versions of the two editions can differ, it's not
    of much use for our purpose.

    To get the content of the `current.ver` file, use:

    ``` r
    # RStudio Desktop
    readr::read_lines(file = "https://download1.rstudio.org/current.ver")
    # RStudio Server
    readr::read_lines(file = "https://download2.rstudio.org/current.ver")
    ```

-   RStudio [provides](https://support.rstudio.com/hc/en-us/articles/203842428-Getting-the-newest-RStudio-builds) canonical URLs for the latest binary releases
    for every supported OS. Examples:

    -   stable:

        -   DEB: <https://rstudio.org/download/latest/stable/desktop/bionic/rstudio-latest-amd64.deb>
        -   RPM: <https://rstudio.org/download/latest/stable/desktop/centos8/rstudio-latest-x86_64.rpm>
        -   EXE: <https://rstudio.org/download/latest/stable/desktop/windows/RStudio-latest.exe>
        -   DMG: <https://rstudio.org/download/latest/stable/desktop/mac/RStudio-latest.dmg>

    -   preview: <https://rstudio.org/download/latest/preview/desktop/bionic/rstudio-latest-amd64.deb>

    -   daily: <https://rstudio.org/download/latest/daily/desktop/bionic/rstudio-latest-amd64.deb>

```{r}
#' Get latest RStudio version number
#'
#' @inheritParams releases
#' @param pro `TRUE` for the proprietary RStudio (Server) Pro edition and `FALSE` for the open-source RStudio (Server) edition.
#' @param os The OS _codename_ for which the RStudio version was built. If `NULL`, it will be auto-detected for the current system.
#'
#' @return `r pkgsnip::return_label("version_nr")`
#' @export
#'
#' @examples
#' latest_version(os = "macos",
#'                use_cache = FALSE)
latest_version <- function(type = c("desktop", "server"),
                           stable = TRUE,
                           pro = FALSE,
                           os = NULL,
                           use_cache = TRUE,
                           max_cache_age = "1 day") {
  
  checkmate::assert_flag(pro)
  
  data <-
    rlang::arg_match(type) %>%
    releases(stable = stable,
             use_cache = use_cache,
             max_cache_age = max_cache_age) %>%
    dplyr::filter(is_pro == pro)
  
  supported_os <-
    data[["os"]] %>%
    unique() %>%
    setdiff(NA)
  
  if (!is.null(os)) {
    
    os <- rlang::arg_match(arg = os,
                           values = supported_os)
  } else {
    
    os <-
      if (xfun::is_linux()) {
        system2(command = "lsb_release",
                args = "-cs",
                stdout = TRUE,
                stderr = TRUE)
      } else if (xfun::is_macos()) {
        "macos"
      } else if (xfun::is_windows()) {
        "windows"
      } else {
        cli::cli_abort("Unknown operating system detected.")
      }
    
    if (!(os %in% supported_os)) {
      
      cli::cli_abort(paste0("The RStudio release suited to your Linux distribution {.field {utils::sessionInfo('base')$running}} codename {.field {os}} ",
                            "couldn't be auto-detected. Please set {.arg os} to one of {.or {.val {supported_os}}}."))
    }
  }
  
  data %>%
    dplyr::filter(os == os) %$%
    version %>%
    max() %>%
    unique() %>%
    as.numeric_version()
}
```

## `releases`

TODO:

-   Improve parsing the XML file containing all the info. [AWS
    `ListObjects`](https://docs.aws.amazon.com/AmazonS3/latest/API/API_ListObjects.html#API_ListObjects_ResponseSyntax) only returns the newest 1'000 results
    and `<IsTruncated>true</IsTruncated>` indicates that there are more -\> Figure out how to fully retrieve this "paginated" XML!

    We can use [ListObjectsV2](https://docs.aws.amazon.com/AmazonS3/latest/API/API_ListObjectsV2.html) on the preview URL:
    <https://s3.amazonaws.com/rstudio-ide-build/?list-type=2>

    (Neither the ListObjects nor the ListObjectsV2 URL params seem to work on the stable URLs)

    Then we can set URL param `continuation-token` to the value of `NextContinuationToken` to access the next page.

    Or even better: Use `delimiter=/` and progress hand over hand by setting `prefix` to the value of `CommonPrefixes`.

    Or *even* better: Use `aws.s3::get_bucket("rstudio-ide-build")`!

    To do the same for the stable release server, we probably first need to figure out the Amazon S3 bucket behind it...?

```{r}
#' Get RStudio release metadata
#'
#' @param type Either `"desktop"` for [RStudio Desktop](https://rstudio.com/products/rstudio/#rstudio-desktop) or `"server"` for
#'   [RStudio Server](https://rstudio.com/products/rstudio/#rstudio-server) release metadata.
#' @param stable Set to `FALSE` to retrieve release metadata of [RStudio preview builds](https://rstudio.com/products/rstudio/download/preview/) instead of
#'   [stable builds](https://rstudio.com/products/rstudio/download/).
#' @param use_cache `r pkgsnip::param_label("use_cache")`
#' @param max_cache_age `r pkgsnip::param_label("max_cache_age")` Defaults to 1 day (24 hours).
#'
#' @return `r pkgsnip::return_label("data")`
#' @export
#'
#' @examples
#' releases(type = "server",
#'          max_cache_age = "1 year 2 months 3 weeks 4 days 5 hours 6 minutes 7 seconds")
releases <- function(type = c("desktop", "server"),
                     stable = TRUE,
                     use_cache = TRUE,
                     max_cache_age = "1 day") {
  
  type <- rlang::arg_match(type)
  checkmate::assert_flag(stable)
  
  pkgpins::with_cache(expr = get_releases(type = type,
                                          stable = stable),
                      pkg = this_pkg,
                      from_fn = "releases",
                      stable,
                      use_cache = use_cache,
                      max_cache_age = max_cache_age)
}

get_releases <- function(type,
                         stable) {
  stable %>%
    ifelse(yes = paste0("https://download", ifelse(type == "desktop", 1L, 2L), ".rstudio.org/"),
           no = paste0("https://s3.amazonaws.com/rstudio-ide-build/")) %>%
    xml2::read_xml() %>%
    xml2::as_list() %>%
    purrr::keep(~ .x[["Name"]] == ifelse(stable,
                                         glue::glue("rstudio-{type}"),
                                         "rstudio-ide-build")) %>%
    purrr::chuck("ListBucketResult") %>%
    purrr::imap(~ {
      if (.y == "Contents") .x else NULL
    }) %>%
    purrr::compact() %>%
    purrr::map_depth(.depth = 2L,
                     .f = unlist) %>%
    purrr::map(tibble::as_tibble) %>%
    purrr::list_rbind() %>%
    dplyr::rename_with(.cols = everything(),
                       .fn = snakecase::to_snake_case) %>%
    dplyr::mutate(last_modified =
                    last_modified %>%
                    clock::naive_time_parse(format = "%Y-%m-%dT%H:%M:%SZ",
                                            precision = "millisecond") %>%
                    clock::time_point_round(precision = "second") %>%
                    clock::as_date_time(zone = "UTC"),
                  is_pro = stringr::str_detect(string = key,
                                               pattern = "-pro-"),
                  os = stringr::str_extract(string = key,
                                            pattern = "(?<=^desktop/)[^/]+(?=/)"),
                  version = stringr::str_extract(string = key,
                                                 pattern = "(?i)(?<=rstudio-((pro|server)-)?)\\d+([\\.-]\\d+)*")) %>%
    dplyr::filter(key != "current.ver")
}
```

## `bundled_pandoc_version`

```{r}
#' Determine version of Pandoc bundled with RStudio
#'
#' @return `r pkgsnip::return_label("version_nr")`
#' @export
bundled_pandoc_version <- function() {
  
  Sys.getenv("RSTUDIO_PANDOC") %>%
    fs::dir_ls(regexp = "pandoc(\\.exe)?$") %>%
    system2(args = "-v",
            stdout = TRUE,
            stderr = TRUE) %>%
    dplyr::first() %>%
    stringr::str_extract("\\d+(\\.\\d+)*") %>%
    as.numeric_version()
}
```

## `pkg_status`

```{r}
#' List RStudio's R package dependencies' installation status
#'
#' @return `r pkgsnip::return_label("data")`
#' @export
pkg_status <- function() {
  
  rstudioapi::getRStudioPackageDependencies() %$%
    pal::is_pkg_installed(pkg = name,
                          min_version = version) %>%
    tibble::enframe(name = "package",
                    value = "is_installed")
}
```

## Miscellaneous

### `pkg_config`

```{r}
#' `r this_pkg` package configuration metadata
#'
#' A [tibble][tibble::tbl_df] with metadata of all possible `r this_pkg` package configuration options. See [pal::pkg_config_val()] for more information.
#'
#' @format `r pkgsnip::return_label("data_cols", cols = colnames(pkg_config))`
#' @export
#'
#' @examples
#' rstd::pkg_config
"pkg_config"
```
